\section{Provenance questions}\label{sup:provquestions}

% \todorenske{How to organize this section? Idea 1: leave it as it is. Idea 2: Make sideways table with 1 question per row, add columns for each use case (U1 - U5) and provenance subcomponent (SC1, SC2, SC3, D1, ...). Indicate for which use case and provenance component each question is applicable. }

\newlist{provquestions}{enumerate}{1}
\setlist[provquestions, 1]
{label=Q\arabic{provquestionsi}., %1., 2., 3., ...
leftmargin=2\parindent,
rightmargin=10pt
}


\subsection*{\ref{uc:wf_dev}: Workflow development}

\begin{provquestions}
    \item What is the influence of including filtering step X on performance? (SC1, SC2, EX1)
    \item What is the influence of a different model architecture on performance? (SC1, SC2, EX1)
    \item What is the influence of a different training set on performance? (SC2, EX1)
    \item What is the influence of a different value for input parameter X on performance? (SC2, EX1)
    \item What is the influence of removing model training feature X on performance? (SC2, EX1)
    \item What is the difference between workflow runs X and Y? (SC3)
    \item Which of workflow runs X and Y used more RAM? (SC3, EX2)
    \item Under which conditions am I allowed to reuse this third-party dataset as input for my workflow? (D3)
    \item Under which conditions am I allowed to reuse this third-party software tool for my workflow? (SW3)
    \item What is the influence of ... on training time? (EX1)
\end{provquestions}

\subsection{\ref{uc:writing}: Publishing the workflow}
\begin{provquestions}[resume]
    \item On which previous research was this research built (conceptually)? (SC1)
    \item Which of my colleagues contributed to the conceptualization of this research? What were their contributions? (SC1, WF1, EX4)
    \item Which cleaning operations were performed on workflow input X? (SC2, D1)
    \item Which were all the datasets I used as inputs and should be cited? (D1)
    \item When was dataset X downloaded from database Y? (D1, D2)
    \item What are all the resources which contributed to this research (and should be cited)? (D1, SW1)
    \item Which software was used in this workflow? (SW1)
    \item I reused a custom, unpublished script made by one of my collaborators. How do I give credit to the original authors? (SW1)
    \item I wrote a custom script based on existing code. How do I give credit to the original authors? (SW1)
    \item I wrote a CWL \emph{CommandLineTool} description for existing software. How do I cite it? (SW1)
    \item I reused an existing CWL CommandLineTool description. How do I give credit to the original authors? (WF1)
    \item Which of my colleagues contributed to this workflow to whom I should give credit and/or propose as co-authors? What were their contributions? (WF1)
    \item From which workflow(s) was this workflow derived? (WF1)
    \item I reused a Dockerfile or Docker container image which was written or built by someone else. How do I give credit to the original authors? (ENV3)
\end{provquestions}

\subsection{\ref{uc:understanding}: Understanding the workflow}
\begin{provquestions}[resume]
    \item What is the goal of this analysis? (SC1)  
    \item Why was this step included? (SC1)
    \item Why was this combination of steps chosen? (SC1)
    \item What is the interpretation of this (intermediate) output?  (SC2)
    \item What was the performance of the model? (SC2)
    \item What was the hypothesis of this workflow run? (SC3)
    \item Why was this set of values chosen as inputs for this workflow execution? (SC3)
    \item Where can I find more information about BioDL dataset? (D1)
    \item Which SAbDab query was used to generate the summary file? (D1)
    \item Which UniProt IDs are part of BioDL? (D1)
    \item Which query was issued to PDB? (D1)
    \item How was this figure from the paper generated? (D1)
    \item For which workflow parameter is this dataset an input? (D4)
    \item Which workflow step produced this output file? (D4)
    \item What was the model architecture? (SW2)
    \item Which related tasks were predicted by the model? (SW2)
    \item How were UniProt IDs mapped to PDB IDs? (WF1)
    \item Which input features were used for the model? (WF2)
    \item What are the values of PSP19 for each amino acid type? (WF2)
    \item What is the format of this output file? (WF2)
    \item How should I cite \emph{cwltool}? (EX3)
    \item Who can I contact for more information about the research? (EX4)
\end{provquestions}

\subsection{\ref{uc:reproducing}: Reproducing the workflow}

\begin{provquestions}[resume]
    \item Which HHBlits reference database was used? Which version? (D1)
    \item What was the last modification date of this input file? (D2)
    \item Where can I download HHBlits reference database? (It was not stored in the RO because of its large size.) (D3)
    \item Which version of HHBlits was used? (SW1)
    \item The CWL \emph{CommandLineTool} description did not describe all parameters of this command-line program. Which values were used for the other parameters? (SW2)
    \item Which format does the PDB batch download script need? (SW2)
    \item Where can I download HHBlits? (SW3)
    \item Does this step need network access? (WF3)
    \item How many CPUs are required for this step, according to the workflow description? (WF3)
    \item How much RAM is necessary for this step, according to the workflow description? (WF3)
    \item Which software (and their versions) was installed on the system? (ENV1) 
    \item Which CPU/GPU was installed? (ENV2)
    \item Original author used Dockerfile. Is the image I built on a rerun the same as the one in the original analysis?  (ENV3)
    \item I pulled a Docker image. Is this the same as which was used in the original analysis? (ENV3)
    \item How long does this step take? (EX1)
    \item When was PDB database queried? (EX1)
    \item When was SAbDab database queried? (EX1)
    \item How much memory was used in the original run? (EX2)
    \item Which version of \emph{cwltool} was used in the original run? (EX3)
    \item Who ran the original workflow and how can I contact them for more information? (EX4)
\end{provquestions}

\subsection{\ref{uc:service}: Model as a web service}

\begin{provquestions}[resume]
    \item What was the query which returned the set of proteins for which I made predictions with the model? (SC2, D1)
    \item Which protein sequences were in the training set? (SC2, D1)
    \item Which proteins had epitope annotations? (SC2, D1)
    \item How should I cite this tool? (WF1)
\end{provquestions}